gender and age in food incentive

### 2024-02-19 outputs/paper/paper.qmd
Added new graphs to the quarto document

In this paper, we will provide a deeper analysis and reproduction of the paper, *Incentives and Unintended Consequences: Spillover Effects in Food Choice* (add reference). Replications of Table 1, Figure 2 and 3 were also completed from the raw data in which the original paper is based. The original study was conducted on 1631 students from various elementary schools in a low-income neighborhood in Chicago, IL from kindergarten to grade 8. The purpose of the study was to determine whether there were spillover effects of students making decisions based on their peers' decisions as well as the incentives given to their peers. The study consisted of two stages. In the first stage (G1), each student was given a card from a deck which offered the choice of cookies or grapes, where the grapes sometimes came with a prize as an incentive to be chosen. In the second stage (G2), each student was given the opportunity to change their choice of food based on the initial decisions of their peers. Students were sat in tables of sizes up to 10 students, in which the proportion of cards in the deck that contained incentivized grapes were 0%, 50%, or 100%. Furthermore, some tables were designated as public treatment, in which students were able to see the incentive status of their peers, and some tables were designated as private treatment, in which students were unable to see the incentive status of their peers.

@@

```{r}
#| echo: false
#| warning: false
#| label: fig-foodChoices
#| fig-cap: ": Students food choices from stage G1 to G2 comparing grape to cookie, cookie to grape, or no change."

@@

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: fig-lunchStatus
#| fig-cap: ": Comparing incentivized students choosing grapes if they are eligible or not eligable for the free lunch program"

@@

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: fig-grade
#| fig-cap: ": Comparing incentivized students choosing grapes or cookies by the students grade"

@@

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: fig-gender
#| fig-cap: ": Comparing incentivized students choosing grapes or cookies by the students gender"

@@

# Calculate proportions of students choosing grapes and cookies for each gender
proportions <- incentivized_students %>%
  group_by(male) %>%
  summarise(
    grapes_proportion = mean(grape2 == 1),
    cookies_proportion = mean(grape2 == 0)
  )

@@

# Create the bar chart
ggplot(proportions_long, aes(x = male, y = proportion, fill = choice)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Gender",
       y = "Proportion",
       fill = "Choice") +
  scale_fill_manual(values = c("grapes_proportion" = '#BDBDBD', "cookies_proportion" = '#424242'),
                    labels = c("Grapes", "Cookies")) +
  ylim(0,1) +
  theme_minimal()

@@

>>>>

In this paper, we will provide a deeper analysis and reproduction of the paper, *Incentives and Unintended Consequences: Spillover Effects in Food Choice* [@angelucci_incentives_2019]. Replications of Table 1, Figure 2 and 3 were also completed from the raw data in which the original paper is based. The original study was conducted on 1631 students from various elementary schools in a low-income neighborhood in Chicago, IL from kindergarten to grade 8. The purpose of the study was to determine whether there were spillover effects of students making decisions based on their peers' decisions as well as the incentives given to their peers. The study consisted of two stages. In the first stage (G1), each student was given a card from a deck which offered the choice of cookies or grapes, where the grapes sometimes came with a prize as an incentive to be chosen. In the second stage (G2), each student was given the opportunity to change their choice of food based on the initial decisions of their peers. Students were sat in tables of sizes up to 10 students, in which the proportion of cards in the deck that contained incentivized grapes were 0%, 50%, or 100%. Furthermore, some tables were designated as public treatment, in which students were able to see the incentive status of their peers, and some tables were designated as private treatment, in which students were unable to see the incentive status of their peers.

@@

```{r}
#| echo: false
#| warning: false
#| label: fig-foodChoices
#| fig-cap: "Students food choices from stage G1 to G2 comparing grape to cookie, cookie to grape, or no change."

@@

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: fig-lunchStatus
#| fig-cap: "Comparing incentivized students choosing grapes if they are eligible or not eligable for the free lunch program"

@@

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: fig-grade
#| fig-cap: "Comparing incentivized students choosing grapes or cookies by the students grade"

@@

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: fig-gender
#| fig-cap: "Comparing incentivized students choosing grapes or cookies by the students gender"

@@

# Calculate proportions of students choosing grapes and cookies for each gender
proportions <- incentivized_students |>
  group_by(male) |>
  summarise(
    grapes_proportion = mean(grape2 == 1),
    cookies_proportion = mean(grape2 == 0)
  )

@@

# Create the bar graph
ggplot(proportions_long, aes(x = male, y = proportion, fill = choice)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Gender",
       y = "Proportion",
       fill = "Choice") +
  scale_fill_manual(values = c("grapes_proportion" = '#BDBDBD', "cookies_proportion" = '#424242'),
                    labels = c("Grapes", "Cookies")) +
  ylim(0,1) +
  theme_minimal()

@@

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: fig-publicIncentive
#| fig-cap: "Comparing the proportion of students in the public group who recieved no incentive and whether or not they stayed with their initil choice of cookie or grape or if they switched."

public_incentives <- food_choices|>
  filter(table_incentive_pr == 0)

num_samples <- length(public_incentives$Stayed)

public_incentives <- public_incentives|>
  group_by(Stayed)|>
  summarise(prop = length(Stayed)/num_samples)

# Create the bar graph
ggplot(public_incentives, aes(x = Stayed, y = prop, fill = Stayed)) +
  geom_col() +
  labs(x = "Choice",
       y = "Proportion") +
  scale_fill_manual(values = c("Changed" = '#BDBDBD', "Stayed Cookie" = '#757575', "Stayed Grape" = '#424242')) +
  ylim(0,0.5) +
  theme_minimal()

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: fig-publicNoIncentive
#| fig-cap: "Comparing the proportion of students in the public group who recieved an incentive and whether or not they stayed with their initil choice of cookie or grape or if they switched."

public_no_incentives <- food_choices|>
  filter(table_incentive_pr == 1)

num_samples <- length(public_no_incentives$Stayed)

public_no_incentives <- public_no_incentives|>
  group_by(Stayed)|>
  summarise(prop = length(Stayed)/num_samples)

# Create the bar graph
ggplot(public_no_incentives, aes(x = Stayed, y = prop, fill = Stayed)) +
  geom_col() +
  labs(x = "Choice",
       y = "Proportion") +
  scale_fill_manual(values = c("Changed" = '#BDBDBD', "Stayed Cookie" = '#757575', "Stayed Grape" = '#424242')) +
  ylim(0,0.5) +
  theme_minimal()
```
